Joint Neighborhood Subgraphs Link Prediction

### 2017-06-30 paper.tex
Update .tex and pdf file
F

\begin{abstract}
Graphs are common data structure used to represent data in which the relation between entities are encoded. A number of graph-based methods and systems have been proposed, varying from biology to recomendation to social network. Most methods have to take partially observed graphs as the input to proceed due to the lack of information. This prevents them from showing promissing results. Link prediction becomes an effective paradigm to solve this problem. However, most existing link prediction methods are based on the transitive manner which cannot effectively exploit graph structures. Here we propose a method that first represents each link between two nodes by a graph composed of their neighborhood subgraphs. We then employ a decompositional graph kernel to measure the similarity between links. An empirical experiment on diferent datasets proves that our proposed method shows the state of the art for link prediction.
\keywords{Link prediction, joint neighborhood subgraphs}
\end{abstract}

\section{Introduction and related work}
A huge amount and variety of relational data are available in a wide range of fields. On the one hand, it provides an opportunity to build systems which automatically extract useful information. On the other hand, it raises challenges for machine learning and data scientists to design high performance predictive systems. A common data struture for representing relational data is graph whose nodes describe for entities and links depict their relations. We can easily see graph-based systems in different domains like biology, chemistry or networking. These systems require the data form used for their input as graphs. However, despite the availability of huge data, most graphs are only partially observed graphs. That means a number of actual links in each graph are absent or missing. Tackling this issue is a necessary condition to obtain high performance graph-based predictive systems. As a consequence, many link prediction methods have been proposed with the aim of recovering the missing links. A graph could be associtated with external information sources, so called "side information". Therefore, a link prediction method attempts to exploit the graph topology and side information to build a the model for predicting the presence or absence of link any between couple of nodes. An expected method is required to make use of both such kind of data sources. In this paper, we only focus on the topological information. Thus we the term link prediction methods to refer to topology-based link prediction methods. Conerning link prediction methods, we can classify into two classes: unsupervised learning and supervised learning.

Unsupervised learning methods compute the scores for node couples based on topological features and directly use these scores to consider non-observed links to be present or absent. In \cite{adamic}, a method, \textit{Adamic-Adar}, is proposed in which a score between two nodes is computed by a weighted sum over their common nodes. Nodes with smaller degree are weighted more comparing to the higher ones. Another method also considers node degree when computing proximity is presented. The probability of a existing link connecting two nodes is proportional to the product of their node degrees. Different from the method in \cite{adamic} which considers the common nodes, a link prediction method, \textit{Katz}, is introduced in \cite{katz} that takes into acount the number of common paths with different lengths between two nodes to measure the proximity between them. The shorter paths are given more weights, meanwhile the longer ones are weighted less. Another method which is a modification of \textit{Katz} notated as Leicht-Holme-Newman is presented in \cite{lhni}. In this method the similarity of a node couple is based on an assumption that two nodes are more similar if they have more similar intermediate nodes. Besides, in \cite{matrix-factorization} they also apply a raw singular value decomposition, SVD, on the adjacency matrix for link prediction. Generally, unsupervised learning methods takes the advantage from their complexity since they do not face with learning process. However, in many case, they do not show impressing results.

Supervised learning link prediction methods are based on a learning process which try to learn a curve that separates present links and absent links relied on their characteristics. This group of methods can also be further grouped into four subsets: feature-based models, graph regularization models, latent class models and latent feature models as presented in \cite{matrix-factorization}. In \cite{nonparametric}, a nonparametric latent feature models for link prediction is introduced in which latent features are extracted by based on an adaptation of Bayesain nonparametric approach. Besides, this method is able to concurrently learn number of latent features and assign which entities have these features. Another method related to latent features is presented in \cite{matrix-factorization}, they propose to use matrix factorization to extract latent features. This method can also take the output of unsupervised methods as its input to proceed. They show impressive results with different use of loss functions: square loss function, log loss function. Compared to unsupervised learning methods, supervised methods tend to show better performance in general. However, they normally face with the high compexity in term of computation and memory consumption.

@@

\section{Method}
In this section, we describe our proposed link prediction method which is based on the graph structure exploitation only.
\subsection{Definitions and notation}
We consider a graph $G=(V,E)$ where $V$ is the set of nodes and $E$ is the set of links. The set $E$ can be devided into a subset of observed links ($O$) and a subset of unobserved links ($U$). It is often that the set $O$ is not a complete set since they are many actual links are missing. Therefore, finding most likely set of links in $U$ to be considered as the complement set of $O$ is an important task which impacts on the performance of graph-based systems. Link prediction is the task that aims at exploiting graph structure to rank links in $U$ from the most to the least probable to be considered as observed links in $O$.

We define the \textit{distance} between two nodes $u$ and $v$, notated as $\mathcal{D}(u,v)$, is the length of the shortest path between them. The \textit{neighborhood} of a node $u$ with radius $r$ is the set of nodes such that they are at distance no greater than $r$ from $u$ and it is denoted by $N_r(u)$ and $N_r(u) = \lbrace v\ |\ d(u,v) \leq r \rbrace$. The \textit{neighborhood subgraph} with radius $r$ of a given node $u$, denoted by $\mathcal{N}_{r}^{u}$, is the subgraph formed by the nodes in the
\textit{neighborhood} with radius $r$ of $u$ and the relative edges with endpoints in $N_r(u)$. The degree of a node $u$ denoted as $d(u)$ is the cardinality of its neighborhood $d(u) = |\mathcal{N}_{1}^{u}|$. We also notate the maximum node degree in the graph $G$ as $d(G)$, $d(G) = max\lbrace d(u)\ |\ u \in V \rbrace$
\subsection{The neighborhood subgraph pairwise distance kernel}
In this section, we briefly desbribe an efficient kernel named NSPDK \cite{nspdk} which allows to measure the similarity between labeled graphs. NSPDK is an instance of decompositional kernel \cite{convolution-kernel}.

@@

\subsection{Node labeling}
NSPDK takes labeled graphs as its input to proceed. Therefore, graph node labeling is a step needs to be done before employing NSPDK for graph similarity measure. Here, we propose to a node labeling fuction $\ell$ such that nodes with similar degrees are assigned the same label. It is due to the assumption that nodes with similar degrees tend to share common properties. Following is the definition of the the function:
\begin{center}
$\ell(u) = \left\{
	\begin{array}{ll}
		d(u),\  & \mbox{if } d(u) \leq 5 \\
		6+i,\ & \mbox{if }  5 + \frac{d(G) - 5}{\lambda-5}*i \leq d(u) < 5 + \frac{d(G) - 5}{\lambda-5}*(i+1)
	\end{array}
\right.$,
\end{center}
where $\lambda > 5$ is the number of labels used to label for nodes in the graphs. The values of $\lambda$ depend on the histogram of graph node degree. We can use tuning technique in order to obtain suitable value for $\lambda$.
\subsection{Link representation}
Most methods for link prediction attempt to computes similarities between nodes and consider to rank links. We desire to take links as examples to learn and predict. Therefore, we propse a method which model the context of two nodes to represent for link connecting them.

@@

\section{Experiment}
In this section, we intend to measure the performance of our proposed method and compare it with other link prediction methods. We follow the procedure used in \cite{matrix-factorization} which employs six datasets in different fields. Following, we give each dataset a short description.
\begin{itemize}
\item \textit{Protein }\cite{protein-protein}: the medium confidence network which encodes the interaction between proteins. It contains 2617 nodes and 11855 links.

@@

\item \textit{Nips} \cite{nips}: a network of co-authors at NIPS conference from first to twelfth edition. Nodes are authors and links encode co-author relation between authors. This network contains 2865 nodes and 4733 links.

>>>>

\begin{abstract}
Graphs are common data structure used to represent data in which the relation between entities are encoded. A number of graph-based methods and systems have been proposed, varying from biology to recommendation to social network. Most methods have to take partially observed graphs as the input to proceed due to the lack of information. This prevents them from showing promising results. Link prediction becomes an effective paradigm to solve this problem. However, most existing link prediction methods are based on the transitive manner which cannot effectively exploit graph structures. Here we propose a method that first represents each link between two nodes by a graph composed of their neighborhood subgraphs. We then employ a decompositional graph kernel to measure the similarity between links. An empirical experiment on different datasets proves that our proposed method shows the state of the art for link prediction.
\keywords{Link prediction, joint neighborhood subgraphs}
\end{abstract}

\section{Introduction and related work}
A huge amount and variety of relational data are available in a wide range of fields. On the one hand, it provides an opportunity to build systems which automatically extract useful information. On the other hand, it raises challenges for machine learning and data scientists to design high performance predictive systems. A common data structure for representing relational data is graph whose nodes describe for entities and links depict their relations. We can easily see graph-based systems in different domains like biology, chemistry or networking. These systems require the data form used for their input as graphs. However, despite the availability of huge data, most graphs are only partially observed graphs. That means a number of actual links in each graph are absent or missing. Tackling this issue is a necessary condition to obtain high performance graph-based predictive systems. As a consequence, many link prediction methods have been proposed with the aim of recovering the missing links. A graph could be associated with external information sources, so called "side information". Therefore, a link prediction method attempts to exploit the graph topology and side information to build a the model for predicting the presence or absence of link any between couple of nodes. An expected method is required to make use of both such kind of data sources. In this paper, we only focus on the topological information. Thus we the term link prediction methods to refer to topology-based link prediction methods. Concerning link prediction methods, we can classify into two classes: unsupervised learning and supervised learning.

Unsupervised learning methods compute the scores for node couples based on topological features and directly use these scores to consider non-observed links to be present or absent. In \cite{adamic}, a method, \textit{Adamic-Adar}, is proposed in which a score between two nodes is computed by a weighted sum over their common nodes. Nodes with smaller degree are weighted more comparing to the higher ones. Another method also considers node degree when computing proximity is presented. The probability of a existing link connecting two nodes is proportional to the product of their node degrees. Different from the method in \cite{adamic} which considers the common nodes, a link prediction method, \textit{Katz}, is introduced in \cite{katz} that takes into account the number of common paths with different lengths between two nodes to measure the proximity between them. The shorter paths are given more weights, meanwhile the longer ones are weighted less. Another method which is a modification of \textit{Katz} notated as Leicht-Holme-Newman is presented in \cite{lhni}. In this method the similarity of a node couple is based on an assumption that two nodes are more similar if they have more similar intermediate nodes. Besides, in \cite{matrix-factorization} they also apply a raw singular value decomposition, SVD, on the adjacency matrix for link prediction. Generally, unsupervised learning methods takes the advantage from their complexity since they do not face with learning process. However, in many case, they do not show impressing results.

Supervised learning link prediction methods are based on a learning process which try to learn a curve that separates present links and absent links relied on their characteristics. This group of methods can also be further grouped into four subsets: feature-based models, graph regularization models, latent class models and latent feature models as presented in \cite{matrix-factorization}. In \cite{nonparametric}, a nonparametric latent feature models for link prediction is introduced in which latent features are extracted by based on an adaptation of Bayesain nonparametric approach. Besides, this method is able to concurrently learn number of latent features and assign which entities have these features. Another method related to latent features is presented in \cite{matrix-factorization}, they propose to use matrix factorization to extract latent features. This method can also take the output of unsupervised methods as its input to proceed. They show impressive results with different use of loss functions: square loss function, log loss function. Compared to unsupervised learning methods, supervised methods tend to show better performance in general. However, they normally face with the high complexity in term of computation and memory consumption.

@@

\section{Method}
In this section, we describe our proposed link prediction method which is based on the graph structure exploitation only.
\subsection{Definitions and notation}
We consider a graph $G=(V,E)$ where $V$ is the set of nodes and $E$ is the set of links. The set $E$ can be divided into a subset of observed links ($O$) and a subset of unobserved links ($U$). It is often that the set $O$ is not a complete set since they are many actual links are missing. Therefore, finding most likely set of links in $U$ to be considered as the complement set of $O$ is an important task which impacts on the performance of graph-based systems. Link prediction is the task that aims at exploiting graph structure to rank links in $U$ from the most to the least probable to be considered as observed links in $O$.

We define the \textit{distance} between two nodes $u$ and $v$, notated as $\mathcal{D}(u,v)$, is the length of the shortest path between them. The \textit{neighborhood} of a node $u$ with radius $r$ is the set of nodes such that they are at distance no greater than $r$ from $u$ and it is denoted by $N_r(u)$ and $N_r(u) = \lbrace v\ |\ d(u,v) \leq r \rbrace$. The \textit{neighborhood subgraph} with radius $r$ of a given node $u$, denoted by $\mathcal{N}_{r}^{u}$, is the subgraph formed by the nodes in the
\textit{neighborhood} with radius $r$ of $u$ and the corresponding edges with endpoints in $N_r(u)$. The degree of a node $u$ denoted as $d(u)$ is the cardinality of its neighborhood $d(u) = |\mathcal{N}_{1}^{u}|$. We also notate the maximum node degree in the graph $G$ as $d(G)$, $d(G) = max\lbrace d(u)\ |\ u \in V \rbrace$
\subsection{The neighborhood subgraph pairwise distance kernel}
In this section, we briefly desbribe an efficient kernel named NSPDK \cite{nspdk} which allows to measure the similarity between labeled graphs. NSPDK is an instance of decompositional kernel \cite{convolution-kernel}.

@@

\subsection{Node labeling}
NSPDK takes labeled graphs as its input to proceed. Therefore, graph node labeling is a step needs to be done before employing NSPDK for graph similarity measure. Here, we propose to a node labeling function $\ell$ such that nodes with similar degrees are assigned the same label. It is due to the assumption that nodes with similar degrees tend to share common properties. Following is the definition of the the function:
\begin{center}
$\ell(u) = \left\{
	\begin{array}{ll}
		d(u),\  & \mbox{if } d(u) \leq 5 \\
		6+i,\ & \mbox{if }  5 + \frac{d(G) - 5}{\lambda-5}*i \leq d(u) < 5 + \frac{d(G) - 5}{\lambda-5}*(i+1)
	\end{array}
\right.$,
\end{center}
where $\lambda > 5$ is the number of labels used to label for nodes in the graphs. The values of $\lambda$ depend on the histogram of graph node degree. We can use tuning technique in order to obtain suitable value for $\lambda$.
\subsection{Link representation}
Most methods for link prediction attempt to computes similarities between nodes and consider to rank links. We desire to take links as examples to learn and predict. Therefore, we propose a method which model the context of two nodes to represent for link connecting them.

@@

\section{Experiment}
In this section, we intend to measure the performance of our proposed method and compare it with other link prediction methods. We follow the procedure used in \cite{matrix-factorization} which employs six datasets in different domains. In the following we shortly describe each dataset.
\begin{itemize}
\item \textit{Protein }\cite{protein-protein}: the medium confidence network which encodes the interaction between proteins. It contains 2617 nodes and 11855 links.

@@

\item \textit{Nips} \cite{nips}: a network of co-authors at NIPS conference from the first to the $12^{th}$ edition. Nodes are authors and links encode co-author relation between authors. This network contains 2865 nodes and 4733 links.
